Doors Open Ottawa 2018

### 2018-07-19 pnsac/newsletter/article_EditorsNotes.tex
Newsletter source file changes for July 2018.

As the President has noted the Association has, through its volunteers,
expended thousands of hours on the North Star's restoration since 2004. In the
last couple of years, however, various developments at CASM have resulted in a
diminution in the amount of restoration work we have been able to do. The result
has been a decline in the number of volunteers working on the aircraft, and a
corresponding decline in our membership. This in turn has led to an overall
decline in the Association's activities, including the publication of this
newsletter. CASM has made some major changes to its personnel, including those
directly responsible for the restoration of the North Star. The Association is
already experiencing a positive change in its relationship with the Museum which
can only be to the good. We look forward to a renewed partnership with CASM to
our mutual benefit.

%\vspace{12pt}

>>>>

Our readers will appreciate from the articles and photographs in this edition that real progress
being made on the restoration of the aircraft. A number of new restoration volunteers have signed on this year bringing with them diverse skills and experience. Whoever thought that someone with a mountain climbing background would be able to put his special skills to work on an aircraft. Believe it or not but see the "Conservator's Corner."  As I have previously commented: When work on the aircraft is moving ahead it encourages enthusiasm at all levels of the organization. One of the biggest cheerleaders of our Association is its president, Richard Lodge. Richard wears many hats including that of a restoration volunteer so he knows of what he speaks.  In this edition Richard discusses how he came to be involved in Project North Star and joins the illustrious members of the Association who have contributed to "Our Members" column.  But I would be remiss if I didn't thank all of our contributors to this edition including our publisher Drew Hodge.

### 2018-07-19 pnsac/newsletter/article_MemberInterview.tex
Newsletter source file changes for July 2018.

\title{Our Members}
\author{Interview with Austin J Timmins}

@@

1.\textit{What is your background in aviation?}

My aviation experience includes 36 years service in the RCAF and nine
years with the Canadian Aeronautics and Space Institute. In the RCAF I
completed radio navigator training, which included communications,
navigation and air gunner. I later qualified as an airborne
interception navigator. My operational experience includes air
transport, search and rescue and training. My assignments included
Navigation Leader, Aircrew Leader, Squadron Commander, Air Command
Navigator and Deputy Commander (COO) of the Air Transport Group. I
logged 9500 hours, mostly on the following aircraft: Beechcraft,
Dakota, North Star, Mitchell, Yukon, Hercules, and Boeing 707. My
staff assignments in Europe included movement control (traffic
management) for 1 Air Division and a tour at The NATO Defense College
in Rome.

On retirement from the RCAF I participated in a Transport
Canada A Base review of several airports. I then secured full
employment with Canadian Aeronautics and Space Institute as Executive
Director, retiring in 1995.

2. \textit{How long have you been involved with Project North Star and how and why did you get involved?}

My involvement with efforts to conserve North Star 17515 predates
Project North Star. Ottawa area members of the 426 Thunderbird Squadron
Association, mostly retired North Star aircrew, met with the National
Aviation Museum staff to express their concern about the state of the
North Star and proposed volunteer involvement in a preservation effort.
These expressions of concern and interest led to a meeting between the
Association representatives, where General Adamson presented a proposal
for the restoration of the aircraft. It was rejected but assurance was
given that the North Star would be first priority when the new hangar
was completed Also a North Star Trust Fund would be established. Robert
Holmgren, an observer at the meeting, then proceeded to develop a
proposal for volunteers to restore the North Star.

@@

3. \textit{What has been the history of your involvement to date?}

I have been involved from the beginning of efforts to establish PNS,
first as a member of a Steering Committee, led by Robert Holmgren,
later, Vice President and then President of PNSAC. My contacts in
aviation provided much needed advice and support, including access to
initial funding for PNSAC.

@@

4. \textit{What has been the highlight of your involvement?}

The highlight of my involvement has to be the establishment of PNS and
PNSAC.

@@

5. \textit{What has been the most challenging part of your involvement?}

PNS is now in its 16th year. It has been a work in progress with many
bumps along the way. I spent a lot of time and effort, along with other
members of the Executive, evening out these bumps. Establishing and
maintaining effective channels of communication with CASM and CSTM
required constant attention. Much effort was directed to fund raising,
publicity, and, finding volunteers. In the beginning, every date
brought a new challenge.

>>>>

\title{Our Members}
\author{Interview with Richard Lodge}

@@

1.\textit{What is your background in aviation?}\\

I trained in England as a Chartered Accountant, now often known as a bean
counter. After I obtained my designation, I wanted to work in industry rather
than in the accountancy profession. Following several interviews, I was offered
a job with the Aero Engine Division of Rolls-Royce Ltd. in Derby, England.

Until I joined Rolls-Royce I had very little interest aviation. I had always
been interested in transportation and had a major passion for railways which was
the mode of transport I normally used in Britain in 1950s and 1960s.

I became interested in aviation at Rolls-Royce and was fascinated by the
complexity of designing and building aero engines.

After leaving Rolls-Royce in 1966, I was not involved with aviation until I came
to Canada in 1978 and returned to the accountancy profession and started doing
tax returns for pilots and other aviation personnel located around Dorval,
Qu\`{e}bec.\\

2. \textit{How long have you been involved with Project North Star and how and why did you get involved?}\\

One of the driving forces behind the foundation of Project North Star was
Robert Holmgren who was among a group who persuaded the National Aviation and
Museum, as it was known then, that a group of a volunteers should be allowed to
work on the North Star and bring it back to Museum display standards. The
Project North Star Association of Canada (PNSAC) was incorporated in 2003 with
Robert as its first president. Robert had worked for Air Canada on the
engineering side of the airline in Dorval and I got to know him when doing the
accounting for his wife's antiques business. At the time of the introduction of
the Québec language laws (Bills 22 and 101), many Anglophones left the Montr\'{e}al
area and moved to Ottawa. Both Robert and I moved in the early 1980's. After
transferring my business to Ottawa, I remained in contact with Robert
professionally and when PNSAC was formed he asked me to become the Association's
first Treasurer.\\

@@

3. \textit{What has been the history of your involvement to date?}\\

I remained Treasurer until 2010 when I was elected President after our second
President, Tim Timmins, stood down. My election as President coincided with my
partial retirement from the accountancy profession. Apart from my interest in
railways, I had also always been interested in engineering and retirement gave
me the opportunity to become a Restoration Volunteer on the North Star. Working
on a Merlin engine and becoming part of the Engine Shop crew was an obvious
place for me to work. I have enjoyed developing some of the limited skills I
learned when working on pre-WW2 cars in England, which mostly used similar
technology to the Merlin engines.\\

@@

4. \textit{What has been the highlight of your involvement?}\\

In the 1950s I was one of the early volunteers working on the first preserved
railway in the world, the Talyllyn Railway in Wales. During this I learned much
about volunteering and how such an organization could work. Helping to organize
and build PNSAC over the last 14 years has both been challenging and very
interesting. We are now seeing the fruits of our work both in the actual
achievements in the restoration of the North Star and the enthusiasm of our
Association members.\\

@@

5. \textit{What has been the most challenging part of your involvement?}\\

Volunteers working on an aircraft owned by the Government of Canada in a
national Federal museum was in the early 2000s an entirely new concept. Until
that time all work on the aircraft collection had been carried out by Museum
permanent staff. It was not easy to persuade the Museum that volunteers could be
trusted to carry out restoration work to the high standards of the Museum and
that as a group they would have the tenacity to continue working for many years.
Now 14 years later and after over 80,000 hours of volunteer restoration work,
PNSAC is going strong and will definitely see the end of the North Star
restoration in the years to come.

Starting a new project is not easy. There is always skepticism as to whether the
new organization will succeed and in some cases resistance to the concept of the
new organization. Over the years we have had to overcome all these problems and
avoid being discouraged when we were unable to do things the way we initially
wanted to do them. Our acceptance by the Museum as a trusted partner is a
testament to the patience and hard work of all our volunteers and the support of
key members of the Museum's permanent staff.

### 2018-07-19 pnsac/newsletter/article_PresReport.tex
Newsletter source file changes for July 2018.

Since the last issue of the NStar Chronicle in March of this year we
have continued to reorganize our volunteer operations under the
leadership of the Museum's Special Projects Manager, R\'{e}jean (Rej)
Demers. We have started a recruitment process for new restoration
volunteers and to date two new people have started work. During the
summer two more permanent Museum staff joined Rej to replace people who
have retired. It is obviously taking time for the new Museum staff to
get to know the complex operations of the Aviation Museum, to get to
know the Project North Star volunteers and assess their various talents
and areas of expertise. Consequently, our summer has been quiet but
steadily productive.

For several reasons it became necessary to reorganize the location of
several major Museum artifacts. This affected the North Star aircraft
which spent the summer outside the Reserve Hanger resulting in some
problems with water ingress and further bird activity. I am pleased to
say that the aircraft is now back inside the hangar.

The Museum is actively supporting Project North Star and trying to
facilitate our restoration work wherever possible and is promoting our
work and future activities. Our Treasurer, Paul Labranche, and I
recently met with the Museum Foundation. The Foundation is the umbrella
organization with responsibility for fundraising and business
development at the three museums, comprising the Museum Corporation,
(Aviation, Science \& Tech and Agriculture). We had a very productive
meeting, which among other things has resulted in the Foundation
deciding to highlight Project North Star in its Fall fundraising mail
out. This type of collaboration is very good for our Association and we
hope as good for the Foundation.

Our biggest challenge at present is to get our website properly
operational again. A couple of our volunteers with website knowledge
have indicated a willingness to do the necessary work and we are hoping
to have the site working and up to date in the near future.

On December 2nd we held a Members' meeting at the Museum when we were
fortunate to have Jim Mason join us to talk about the restoration of an
early Lockheed aircraft flown by TCA. The meeting was well attended in
a new venue of the Avro Arrow room instead of the Bush Theatre.
Following Jim's interesting presentation, the group moved to the
Museum's Restoration Hangar where Rej talked about the progress of the
North Star restoration and the plans for work in 2018.

One of the initiatives currently underway is to create a visual
timeline of the work already completed on the North Star restoration
and the planned future work to bring the project to completion. It is
anticipated that this visual timeline will be mounted on a display
board for our members and the public to view.

In closing, I am optimistic that 2018 will see increased activity both
in the aircraft restoration work and also within our Association.

>>>>

These are exciting times to be a member of the Project North Star Association.
We  have recently taken part in the 2018 Doors Open Ottawa. This is a major
event for the Museum and PNSAC volunteers were able to make a significant
contribution to the success of the two open days. Doors Open has become second
only to Canada Day in its importance to the Museum. (more on Open Doors 2018
later in this issue).

A few days before Canada Day the aircraft was brought outside. Unfortunately,
the day itself was extremely hot and the number of visitors was well down from
previous years. Nonetheless our volunteers did a yeoman service in talking to
the public about our restoration work.

Although showing the aircraft to the public is very important to us, our main
purpose is to restore the North Star to museum condition. Much of our work is
detailed, painstaking and unspectacular such as restoring a small part of the
aircraft which can take many days or weeks. At other times making progress is
very visible. We are in such a situation now.

At the end of May we had one of our regular meetings with Chris Kitzan (Director
General of CASM) together with Neil Raynor (VP of PNSAC), Cedric St-Amour (CASM
Volunteer Coordinator) and Reg Demers (Our CASM supervisor There was a full
exchange of ideas, suggestions for bringing  the public into the Museum and ways
in which PNSAC could be involved in these initiatives and events. The meeting
ended with me having a full sheet of notes covering future planning, major
restoration work and how we can make sure we recruit younger people both as
members of PNSAC and where possible active Restoration Volunteers.

A meeting like the one above only happens when an organization is thriving, and
everybody is excited about the future. During the last 15 years PNSAC has
demonstrated that it is capable delivering very high-quality restoration work.
We have already accumulated over 80,000 hours of work and during this time we
have become a trusted partner of the Museum. Our website is now operational
again and in the next few weeks we expect to be able to update it and make it
more relevant to where we are with the project now. One of the major initiatives
taking place is to develop a diagrammatic history of the work done on the North
Star restoration and a timeline for future work on the aircraft.

In closing, I once again urge Association members to become active volunteers.
There are many volunteering opportunities both for those with major aviation
experience to those who can only volunteer with us on occasions like Canada Day
or can provide help for our many behind-the-scenes activities.

Our next Members' Meeting is scheduled for
Saturday, September 29, 2018. Do reserve this date in your calendar if you can
be in the Ottawa area on that day. The meeting will be held at the Canada
Aviation and Space Museum. More details will follow later.

### 2018-07-19 pnsac/newsletter/article_ProgReport.tex
Newsletter source file changes for July 2018.

\title{Conservator's Corner}
%\subtitle{PNS 2016 Status Update}
\author{R\'{e}jean Demers}

@@

\textit{R\'{e}jean Demers (Rej, to those who work with him) is a Transport Canada
licensed Aircraft Maintenance Engineer (AME). R\'{e}jean was a small
business owner in Gatineau QC for over 10 years, working for a wide
variety of private and commercial aircraft owners and operators.
Following a stint in Northern Ontario working on turpoprop freighters
and bulk fuel haulers, he took up his present position as a
Conservator, Special Projects Manager with Ingenium, Canada's Museums
of Science and Innovation in January 2017. His main project is the
restoration of the Canadair North Star DC-4GM at the Canadian Aviation
and Space Museum. He recounts below his introduction to his new job.}

Leaving Pickle Lake was like getting shot out of a cannon in a snow
storm. At that time, Christmas of 2016, I was the last person on my
crew to leave the YPL base, having seen my colleagues off for the end
of our tour and the holidays. After a full day of flying, the fleet had
to be tucked away in the barn ready for service. On Boxing Day, a fresh
crew would fly in and start another two week tour of line maintenance
with the company we worked for. However, on that Christmas, I was
quitting my job.

Nobody knew I was leaving until I got back to the bunkhouse and shared
Christmas dinner with the remaining pilots and ops personnel. I refer
to them in these terms now, but then and there, I could say they were
my friends. In the weeks leading to my departure, no one knew I was
leaving. No one speaks of breaking out of prison when they are behind
bars. The difference being the only bars I saw were the Pickle Lake
Hotel and a watering hole in Thunder Bay called the Apollo.

Driving down the 599 in that snow storm with a truck I had bought for
\$100 was a white knuckle ride. Knowing this would probably be the last
time I drove the 2000km distance between Northern Ontario and home in
Gatineau. The stakes were high: A flight to catch in Ottawa (T-minus
48hrs), New Year's eve in New York city with my family waiting there
for me and a new job starting at CASM on the day of my wedding
anniversary. Could 2017 start with anything less? Arriving to work that
January 3rd morning was a good feeling, until I got startled by "Manny"
the mannequin sitting in my predecessor's office chair. If a sense of
humour was a job requirement, I hoped to get the joke. I would later
find Manny in other compromising positions, some of which I refuse to
admit my participation in writing.

The momentum I have gained in the twelve months since my escape from
Pickle Lake is only part of what keeps me moving forward. I also count
on the support of others, namely co-workers and volunteers. My position
as Project Manager of the North Star restoration relies on the good
work and cooperation of all hands involved. As the work itself proves
to be technical and highly detailed, with wide ranges of abilities
involved, people are always the driving force.

It was suggested my article for the North Star Chronicle be focused on
the restoration of a certain four engined artifact \#1967.0645. For
details on the work that has been carried out by our dedicated group of
volunteers, I recommend you read my conservation report on the subject.
It is available online: \textcolor{blue}{\texttt{\url{http://www.projectnorthstar.ca}}}
or by email: \textcolor{blue}{\email{rdemers@techno-science.ca}}

>>>>

\title{Conservator's Corner}
%\subtitle{PNS 2016 Status Update}
\author{R\'{e}jean Demers\\ Conservator\\ Canadian Aviation and Space Museum}

@@

\textbf{Stabs and Tail Feathers}

Some tasks are easier than others, this is not a simple one. North Star S/N:
122, registration 17515, was delivered to the National Aeronautical collection
in 1965, intact and airworthy. After spending 39 years parked on the tarmac,
awaiting restoration, the condition of its flight control surfaces had become
less a statement of endurance and more a sad result of the elements.

\begin{figure}[httb]
   \vspace{2em}
   \centering
   \includegraphics[scale=0.5]{IMG_0705-scaled.png}
   \caption*{\small \em Scaffold erected next to North Star to support removal of the rudder and vertical stablizer.}
   \label{fig:stab-one}
\end{figure}

By the time restoration had begun in 2004, the flight control surfaces had
sustained the full range of what mother nature had to offer over the years. In
the fourteen years that followed, work has been carried out removing what
assemblies could be reached from the ground. Ailerons and flaps, elevators and
horizontal stabilizers had been removed and now, the vertical stabilizer and
rudder in 2018.

This article will focus on the year long process that occupied roughly 2-4
volunteers, twice a week, working at heights on scaffolding around the empennage
("tail assembly") of 515. Work commenced following the museum's last big
shuffle, in the fall of 2017. One side of the storage hangar was emptied to
accommodate the acquisition of the C130E Hercules and Convair 580. The removal
of CP-107 Canadair Argus from storage was a necessary trade-off for the new
stacking, that left Project North Star with just enough space to work on.

%\begin{figure}[httb]
%   \vspace{2em}
%   \centering
%   \includegraphics[scale=1.25]{DSC_1156-scaled.png}
%   \caption*{\small \em \color{red} CAPTION NEEDED}
%   \label{fig:stab-one}
%\end{figure}

The work involved in removing flight control surfaces from the ship starts by
disconnecting cables, push rods and torque tubes. These relay pilot input
directly to the airfoil. There are secondary components such as panels,
fairings, plumbing for de-icing boots, grounding straps and electrical conduits.
Finally, when the control surface or stabiliser is ready to be pulled, hinge
bearings and spar bolts must be freed. Support must be provided to the component
once the weight of the assembly is relieved from the airframe. Once
disconnected, the "tail feather" or "stab" is lowered into a cradle on the
ground for cleaning, condition assessment and preparation for restoration. The
approach being that a component is more accessible at ground level than 10-20
feet overhead. Eventually, the outboard wing sections will follow the same
treatment: Not a light undertaking.

Volunteers on task are comprised of a mix of experienced members and new
additions to the crew. Setting up the scaffold was the first requirement in
gaining access. We had an ex roughneck, Will Assad, building the first tower.
Chris McGuffin, an experienced mountaineer, built the second stage to a full
height of 30ft. Bruce Gemmill donated blood removing the fin retaining hardware,
a modest 64 bolts through 4in. access panels. Peter Trobridge spent some hours
in the hell hole, fitting a tail skid extension jack and running leaders for
control cables. The heat of lights and summer humidity had a therapeutic effect,
much like a sauna.

Years of battering had taken its toll on the rudder, shearing the torque tube
and stripping the rudder of it's fabric. During this time, generations of
starlings had been introduced to multi-level condo dwelling. Three garbage bags
of nesting material were removed just to gain access to the pulleys in the
vertical stabiliser alone. Starlings are neither good nest makers nor parents to
their young. I will spare you the graphic descriptions of what was removed from
those access panels.

\begin{figure}[httb]
   \vspace{2em}
   \centering
   \includegraphics[scale=0.55]{IMG_20180529_132544212-scaled.png}
   \caption*{\small \em Volunteer Bruce Gemmill removing bird nests from the vertical stabilizer.}
   \label{fig:stab-one}
\end{figure}

The horizontal stabilisers were removed earlier, yet there still remained some
nesting material to clear. These assemblies were brought to the restoration
shops for a thorough clean-out and condition assessment. Phil Chrysler, Bill
Cole,  Richard Houle and John Makadi lent their hands at pulling more stacks of
hay out of the two structures. Access proved difficult necessitating  the use of
a hinge rods, in the form of hooked tools for retrieving deep tufts of clumped
nesting. The individual stabilisers were flipped over, blown out, pressure
washed, and oiled, eventually rendering an accessible, stable assembly ready for
restoration. We discovered a large patch on the leading edge of the port
stabiliser. It covered a blunt force impact form which was probably caused by
the boom of a forklift in a former life.

At the time of going to press, the rudder has been removed from the aircraft and
the vertical stabiliser is slung and ready to pull. The North Star is prepped to
roll out of storage for Canada day and the vertical stab removal is planned for
July 5th. A 30T crane has been rented for the job, with all hands on deck
confirmed for that day's shift schedule. Following this important milestone, the
maple leaf will be lowered from North Star 515, until such time as the
restoration nears its completion. I hope to see the day, dear reader, when the
maple leaf is flown from those heights again.

### 2018-07-19 pnsac/newsletter/article_SpecialEvents.tex
Source files for Dec13 newsletter.
Newsletter source file changes for July 2018.

\title{Quarterly Meeting Report}
\author{Drew Hodge}

@@

On Saturday 7th December 2013, Project North Star Association of
Canada (PNSAC) held a quarterly meeting in the auditorium of the
Canadian Aviation and Space Museum (CASM) in Ottawa. It was a
memorable meeting with over 200 people in attendance. Those in the
audience that day were able to hear the story of the Gimli Glider told
in person by Captain (retd.) Robert (Bob) Pearson, the pilot of Air
Canada Boeing 767-200, fin number 604, when it made an unscheduled
stop at Gimli Manitoba, thirty years ago.

The meeting started with an introduction by Bill Tate, Project North
Star's Vice President.  Bill began by welcoming the audience, which,
as well as PNSAC volunteers, included members of Vintage Wings of
Canada, Ottawa Airport Watch, the Canadian Aviation Historical
Society, the College of Professional Pilots of Canada, CASM volunteers
and staff, Air Canada and WestJet pilots, and visitors from the
general public, one of whom had travelled to Ottawa from Vancouver for
the meeting. After describing the goals of Project North Star and
asking PNSAC volunteers to stand and be recognized, Bill invited the
audience to visit the world class exhibits in the Museum after the
meeting.

Bill's acquaintance with Bob Pearson goes back thirty-five years, when
Bill was Captain Pearson's Second Officer on Boeing 727s based in
Montreal. His duties in the cockpit back then gave Bill ample
opportunities to watch Captain Pearson at work. Bill described Captain
Pearson as a "highly professional pilot" who, "wearing his command
lightly", demonstrated "highly polished interpersonal skills and a
delightful sense of humour".  He then invited Captain Pearson to tell
the story of flying unscheduled into Gimli.

Captain Pearson spoke for a little over an hour with no notes and with
reference to a few appropriate slides.  He described eloquently how,
because of errors in the conversion of metric measurements, his Boeing
767-200 ran out of fuel and lost both engines over Manitoba on the way
from Ottawa to Edmonton. The full story is well documented in the book
"Freefall: From 41,000 feet to zero - a true story, William and
Marilyn Hoffer, Simon \& Schuster, 1989", but to hear it told in
person by the pilot in command that Saturday, 23rd July, 1983, was a
rare privilege.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{Pearson004.eps}
   %caption of the figure 
   \caption*{\small \em Captain (retd.) Robert Pearson.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:pearson004}
\end{figure}

The port engine stopped first at 41,000 feet prompting the decision to
divert to Winnipeg.  At 28,000 feet the starboard engine ran down too,
and Maurice Quintal, Captain Pearson's First Officer, made
calculations that showed they couldn't reach Winnipeg.  Quintal
advised Captain Pearson to try for the nearest runway on the former
RCAF aerodrome at Gimili, twelve miles away.  Some eight or ten
minutes after losing the first engine, Captain Pearson side-slipped
the big Boeing down to an old runway that was being used as a racing
track by members of the Winnipeg Sports Car Club. The nose wheel
collapsed on landing, but the aircraft came to a stop and no one was
seriously injured. The racers helped to put out a minor fire in the
nose.

Captain Pearson talked about the inquiry that followed the incident
and its eventual exoneration of both pilots.  The aircraft was flown
to Winnipeg just two days after the Gimli landing and saw service with
Air Canada for another twenty-five years.  When the Boeing was retired
to the Mohave desert in 2008, Captain Pearson was at the controls to
make that last landing in Arizona.

Following his talk, Captain Pearson answered questions from the
audience.  Richard Lodge, PNSAC President, thanked Captain Pearson and
presented him with a PNSAC golf shirt, a baseball cap, a coffee mug,
and finally a certificate conferring upon him an Honorary Life
Membership of the Project North Star Association.  Captain Pearson
then graciously agreed to make presentations of volunteer-hours
certificates and 10th Anniversary photos to PNSAC volunteers.  The
following volunteers received certificates for the hours shown:

\begin{itemize}
  \item Bruce Gemmill (7000 hrs)
  \item Ted Devey (4000 hrs)
  \item Bill Tate (3000 hrs)
  \item Richard Lodge (1000 hrs)
  \item Bruce Grant (1000 hrs)
  \item Robert D\'{e}sjardins (1000 hrs)
  \item Peter Trowbridge (1000 hrs)
  \item Charles Baril (1000 hrs)
\end{itemize}

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{Pearson007.eps}
   %caption of the figure 
   \caption*{\small \em Richard Lodge presenting Captain Pearon with a
     certificate for Honorary Life Membership in PNSAC.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:engine3propcowl}
\end{figure}

To conclude the meeting, Captain Pearson presented signed copies of
the book "Freefall: From 41,000 feet to zero - a true story" to three
audience members whose ticket numbers he had drawn at random.

The Association's sincere thanks go to Bill Tate for organizing this
quarterly meeting and the talk by Captain Pearson -- it was largely
Bill's efforts that made the meeting the great success that it was.

>>>>

\title{Doors Open Ottawa 2018}
\author{Roger Button}

@@

Doors Open Ottawa gives the general public an opportunity to view facilities for free which they would
normally not have access. Evidently, it is the second largest event of its kind in North America and one of
the largest in the world with over one million visitors since its inception in 2002. In the case of the
Canadian Aviation and Space Museum its participation meant opening the doors of the reserve hangar.
On June 2nd and 3rd the Museum welcomed some 3000 visitors. The visits of the reserve hangar took
place by way of a self-guided tour supported by many of the Museum's volunteers coordinated by Cedric
St. Amour the volunteer coordinator for the Museums Corporation. Project North Star was ably
represented by Charles Baril, Claire Cameron, Garry Dupont, Ted Devey, Chris McGuffin, Neil Raynor,
Jacques Roy, Nelson Smith, Tim Timmins, and Gary Whitten who ably answered questions about the
aircraft. Rejean Demers, the Conservator Special Projects Manager was also present.

\begin{figure}[httb]
   \vspace{2em}
   \centering
   \includegraphics[scale=0.275]{"PNSAC DOO 2018-scaled".png}
  \caption*{\small \em Volunteers Roger Button, Richard Lodge, Chris McGuffin, Charles Baril and Garry Dupont on hand to explain our project at Doors Open Ottawa June 3rd.}
   \label{fig:stab-one}
\end{figure}

This event is also an opportunity to do a little fundraising for the Association. The merchandise desk was
managed by Richard Lodge, Drew Hodge, and Roger Button. Sales of memorabilia and clothing were
steady throughout the two days. A few memberships were also sold and donations received. The
merchandise table was well positioned near the exit to catch visitors as they left. Unfortunately, the
aircraft was not next to the merchandise table and it was clear that some visitors did not make the
connection and others left without seeing the aircraft. It is hoped that these issues will be addressed for
future viewings of the aircraft. Thanks to all the volunteers who devoted their time to this successful
event.

%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.9]{Pearson004.eps}
%   %caption of the figure 
%   \caption*{\small \em Captain (retd.) Robert Pearson.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:pearson004}
%\end{figure}
%
%
%
%\begin{figure}[htbp]
%   \vspace{2em}
%   \centering
%   %name of the graphic, without the path AND in EPS format:
%   \includegraphics[scale=0.9]{Pearson007.eps}
%   %caption of the figure 
%   \caption*{\small \em Richard Lodge presenting Captain Pearson with a
%     certificate for Honorary Life Membership in PNSAC.}
%   %label of the figure, which has to correspond to \ref{}:
%   \label{fig:engine3propcowl}
%\end{figure}
